---
title: "Lab 13: Logistic Regression"
author: "Name here"
output: pdf_document
---


For this question, we will use a historical data set with NCAA tournament results. However, we will now look at predicting a binary outcome, win or loss by the higher seeded team.

```{r, message = F}
library(rjags)
library(knitr)
library(tidyverse)
ncaa <- read_csv('https://raw.githubusercontent.com/stat456/labs/main/Lab12_data.csv') %>%
  filter(Seed.Diff != 0) %>%
  mutate(Seed.Diff = -1 * Seed.Diff,
         SAG.Diff = -1 * SAG.Diff,
         upset = as.numeric(Result == 'Loss'))

ncaa_train <- ncaa %>% filter(Season < 2020)
ncaa_test <- ncaa %>% filter(Season >= 2020)
```

#### 1. (4 points)

Create two figures to explore the impact of `Seed.Diff` and `SAG.Diff` on `upset.` Add a smoother line to approximate the relationship.


#### 2. (4 points)
Write a short caption to accompany each figure.

#### 3. (4 points)
Deviance Information Criteria (DIC) is a Bayesian analog to AIC. Consider the two model below, which do you prefer and why?

```{r}
# Model String
logistic = "model {
  for ( i in 1:N ) {
    y[i] ~ dbern(p[i])
    logit(p[i]) = beta0 + beta1 * x[i]
  }
  beta0 ~ dnorm(M0, 1 / S0^2)
  beta1 ~ dnorm(M1, 1 / S1^2)
} "
writeLines( logistic, con='logistic.txt')
```

```{r}
# Runs JAGS Model: Seeds
seeds_model <- jags.model( file = "logistic.txt", 
                           data =  list(y = ncaa_train$upset, 
                                        x = ncaa_train$Seed.Diff,
                                        N = nrow(ncaa_train), 
                                        M0 = 0, 
                                        S0 = .001,
                                        M1 = 0, 
                                        S1 = 3),
                         n.chains = 2, n.adapt = 5000)
update(seeds_model, n.iter = 5000)

seeds_coda <- coda.samples(seeds_model, 
                          variable.names = c('beta0','beta1'), 
                          n.iter = 5000)

summary(seeds_coda)

summary(glm(upset ~ Seed.Diff -1, family = binomial, data = ncaa_train))

dic_seed <- dic.samples(seeds_model, 5000)
dic_seed
```

```{r}
# Runs JAGS Model: Sagarin
sag_model <- jags.model( file = "logistic.txt", 
                           data =  list(y = ncaa_train$upset, 
                                        x = ncaa_train$SAG.Diff,
                                        N = nrow(ncaa_train), 
                                        M0 = 0, 
                                        S0 = .001,
                                        M1 = 0, 
                                        S1 = 3),
                         n.chains = 2, n.adapt = 5000)
update(sag_model, n.iter = 5000)

sag_coda <- coda.samples(sag_model, 
                          variable.names = c('beta0','beta1'), 
                          n.iter = 5000)

summary(sag_coda)

dic_sag <- dic.samples(sag_model, 5000)
dic_sag
```

```{r}
diffdic(dic_seed, dic_sag)
```


#### 4. (4 points)

Fit the best possible model using DIC as your criteria. You may want to consider interactions and/or non-linearity terms.


#### 5. (4 points)

Write out the formal model you've selected in part 4, including all priors.




#### 6. (4 points)

Interpret your parameters in the model and summarize your findings. Assume you are telling your parents how statistics can be useful for filling out an NCAA bracket.


#### 7. (4 points)

Construct a posterior predictive distribution to calculate the winning probability of the following games, Note your model is likely set to predict the point spread for the higher seed:

1. Montana State (Seed: 14, Sagarin: 133) vs. Kansas State (Seed: 3, Sagarin: 18)
2. Purdue (Seed: 1, Sagarin: 9) vs. Farleigh Dickinson (Seed: 16, Sagarin: 310)
3. Arizona (Seed: 2, Sagarin: 10) vs. Princeton (Seed 15:, Sagarin: 118)
4. Connecticut (Seed: 4, Sagarin: 4) vs. Iona (Seed: 13, Sagarin: 86)

Hint you might need to use `ilogit()` in JAGS.

#### 8. (1 EC point)

Rather than using DIC as a model selection criteria, compare your model's ability to predict outcomes during the 2021 and 2022 tournaments (`ncca_test`).

Consider using classification error (win/loss) and Brier score (`brier.scoer()` in the `iterativeBMA` package. )
