---
title: "Lab 12: Regression"
author: "Name here"
output: pdf_document
---


For this question, we will use a historical data set with NCAA tournament results.

```{r}
library(rjags)
library(knitr)
library(tidyverse)
ncaa <- read_csv('https://raw.githubusercontent.com/stat456/labs/main/Lab12_data.csv') %>%
  filter(Seed.Diff != 0) %>%
  mutate(Seed.Diff = -1 * Seed.Diff,
         SAG.Diff = -1 * SAG.Diff)
```

#### 1. (4 points)

Create two figures to explore the impact of `Seed.Diff` and `SAG.Diff` on `Score.Diff.` Add a smoother line to approximate the relationship.

```{r, fig.cap = 'The higher seed would, unsurprisingly, be expected to win. We do see some potential evidence of non-linearity.'}
ncaa %>% ggplot(aes(y = Score.Diff, x = Seed.Diff)) + 
  geom_point(alpha = .2) + theme_bw() +
  geom_smooth(method = 'loess', formula = 'y ~ x') +
  ggtitle('Point differential vs. Seed difference from historical NCAA basketball')
```

```{r, fig.cap = 'The negative values correspond to lower seeded teams having higher Sagarin ratings. In general the Sagarin ratings seems to do a good job of estimating score differential - even when ratings do not agree with seeds.'}
ncaa %>% ggplot(aes(y = Score.Diff, x = SAG.Diff)) + 
  geom_point(alpha = .2) + theme_bw() +
  geom_smooth(method = 'loess', formula = 'y ~ x') +
  ggtitle('Point differential vs. Sagarin difference from historical NCAA basketball')

```


#### 2. (4 points)
Write a short caption to accompany each figure.

Added above

#### 3. (4 points)
Deviance Information Criteria (DIC) is a Bayesian analog to AIC. Consider the two model below, which do you prefer and why?

```{r}
# Model String
modelString = "model {
  for ( i in 1:N ) {
    y[i] ~ dnorm(beta0 + beta1 * x[i], 1/sigma^2) # sampling model
  }
  beta0 ~ dnorm(M0,1/S0^2)
  beta1 ~ dnorm(M1, 1 / S1^2)
  sigma ~ dunif(0,C)
} "
writeLines( modelString, con='NORMmodel.txt')
```

```{r}
# Runs JAGS Model: Seeds
seeds_model <- jags.model( file = "NORMmodel.txt", 
                           data =  list(y = ncaa$Score.Diff, 
                                        x = ncaa$Seed.Diff,
                                        N = nrow(ncaa), 
                                        M0 = 0, 
                                        S0 = .001,
                                        M1 = 1, 
                                        S1 = 3, 
                                        C = 100),
                         n.chains = 2, n.adapt = 1000)
update(seeds_model, n.iter = 1000)

seeds_coda <- coda.samples(seeds_model, 
                          variable.names = c('beta0','beta1', 'sigma'), 
                          n.iter = 5000)

summary(seeds_coda)

dic_seed <- dic.samples(seeds_model, 5000)
dic_seed
```

```{r}
# Runs JAGS Model: Sagarin
sag_model <- jags.model( file = "NORMmodel.txt", 
                           data =  list(y = ncaa$Score.Diff, 
                                        x = ncaa$SAG.Diff,
                                        N = nrow(ncaa), 
                                        M0 = 0, 
                                        S0 = .001,
                                        M1 = .5, 
                                        S1 = 3, 
                                        C = 100),
                         n.chains = 2, n.adapt = 1000)
update(sag_model, n.iter = 1000)

sag_coda <- coda.samples(sag_model, 
                          variable.names = c('beta0','beta1', 'sigma'), 
                          n.iter = 5000)

summary(sag_coda)

dic_sag <- dic.samples(sag_model, 5000)
dic_sag
```

```{r}
diffdic(dic_seed, dic_sag)
```

**_DIC suggests that the Sagarin model is superior._**

#### 4. (4 points)

Fit the best possible model using DIC as your criteria. You may want to consider interactions and/or non-linearity terms.

```{r}
# Model String
modelString2 = "model {
  for ( i in 1:N ) {
    y[i] ~ dnorm(beta0 + beta1 * x[i] + beta2 * x2[i], 1/sigma^2) # sampling model
  }
  beta0 ~ dnorm(M0,1/S0^2)
  beta1 ~ dnorm(M1, 1 / S1^2)
  beta2 ~ dnorm(M2, 1 / S2^2)
  sigma ~ dunif(0,C)
} "
writeLines( modelString2, con='NORMmodel2.txt')

# Runs JAGS Model: Sagarin
sag2_model <- jags.model( file = "NORMmodel2.txt", 
                           data =  list(y = ncaa$Score.Diff, 
                                        x = ncaa$SAG.Diff,
                                        x2 = ncaa$Higher.SAG,
                                        N = nrow(ncaa), 
                                        M0 = 0, 
                                        S0 = .001,
                                        M1 = .5, 
                                        S1 = 3,
                                        M2 = 0, 
                                        S2 = 3,
                                        C = 100),
                         n.chains = 2, n.adapt = 1000)
update(sag2_model, n.iter = 1000)

sag2_coda <- coda.samples(sag2_model, 
                          variable.names = c('beta0','beta1','beta2', 'sigma'), 
                          n.iter = 5000)

summary(sag2_coda)

dic_sag2 <- dic.samples(sag2_model, 5000)
dic_sag2
```


#### 5. (4 points)

Write out the formal model you've selected in part 4, including all priors.

\begin{eqnarray}
  pointdiff &= \beta_0 + \beta_1 * x_{sag diff} + \epsilon \;\; \epsilon \sim N(0,\sigma^2)\\
  \beta_0 &\sim N(0, .001^2) \\
  \beta_1 &\sim N(1, 3^2)\\
  \sigma &\sim Unif(0, 100)
\end{eqnarray}


#### 6. (4 points)

Interpret your parameters in the model and summarize your findings. Assume you are telling your parents how statistics can be useful for filling out an NCAA bracket.

__Given Sagarin ratings, seeds are not particularly useful. For each difference in Sagarin ratings, we'd expect to see about a .13 (.12, .14) expected point differential. So for two teams separated by 10 values in the ratings, the lower Sagarin rating team would be expected to win by about 1.3 points. However, there is quite a bit of randomness present, where actual spreads could differ by about 20 points in either direction.__

#### 7. (4 points)

Construct a posterior predictive distribution to calculate the winning probability of the following games, Note your model is likely set to predict the point spread for the higher seed:

1. Montana State (Seed: 14, Sagarin: 133) vs. Kansas State (Seed: 3, Sagarin: 18)
2. Purdue (Seed: 1, Sagarin: 9) vs. Farleigh Dickinson (Seed: 16, Sagarin: 310)
3. Arizona (Seed: 2, Sagarin: 10) vs. Princeton (Seed 15:, Sagarin: 118)
4. Connecticut (Seed: 4, Sagarin: 4) vs. Iona (Seed: 13, Sagarin: 86)

```{r}
# Model String
modelString_pp = "model {
  for ( i in 1:N ) {
    y[i] ~ dnorm(beta0 + beta1 * x[i], 1/sigma^2) # sampling model
  }
  pp_msu ~ dnorm(beta0 + beta1 * 115, 1/sigma^2)
  pp_purdue ~ dnorm(beta0 + beta1 * 301, 1/sigma^2)
  pp_arizona ~ dnorm(beta0 + beta1 * 108, 1/sigma^2)
  pp_uconn ~ dnorm(beta0 + beta1 * 82, 1/sigma^2)
  
  beta0 ~ dnorm(M0,1/S0^2)
  beta1 ~ dnorm(M1, 1 / S1^2)
  sigma ~ dunif(0,C)
} "
writeLines( modelString_pp, con='NORMmodel_pp.txt')
```


```{r}
# Runs JAGS Model: Sagarin
sag_model_pp <- jags.model( file = "NORMmodel_pp.txt", 
                           data =  list(y = ncaa$Score.Diff, 
                                        x = ncaa$SAG.Diff,
                                        N = nrow(ncaa), 
                                        M0 = 0, 
                                        S0 = .001,
                                        M1 = .5, 
                                        S1 = 3, 
                                        C = 100),
                         n.chains = 2, n.adapt = 1000)
update(sag_model_pp, n.iter = 1000)

sag_coda_pp <- coda.samples(sag_model_pp, 
                          variable.names = c('beta0','beta1', 'sigma',
                                             'pp_msu','pp_purdue','pp_arizona','pp_uconn'), 
                          n.iter = 5000)

pp_out <- tibble(vals = c(sag_coda_pp[[1]][,'pp_msu'],
                          sag_coda_pp[[1]][,'pp_purdue'],
                          sag_coda_pp[[1]][,'pp_arizona'],
                          sag_coda_pp[[1]][,'pp_uconn']),
                 type = rep(c('Kansas St', 
                              'Purdue', 
                              'Arizona',
                              'UConn'),
                            each = 5000))

pp_out %>% group_by(type) %>% 
  summarize(`upset prob` = mean(vals < 0)) %>%
  kable(digits = 3)
```

