---
title: "Lab 4: Key"
format: pdf
editor: source
---

Use a dataset containing homes in the Seattle, WA area [http://www.math.montana.edu/ahoegh/teaching/stat408/datasets/SeattleHousing.csv](http://www.math.montana.edu/ahoegh/teaching/stat408/datasets/SeattleHousing.csv) for this question.


Estimate the posterior distribution for the probability that houses in Seattle have more than 2 bathrooms.
```{r, message = F}
library(tidyverse)
library(scales)
seattle <- read_csv('http://www.math.montana.edu/ahoegh/teaching/stat408/datasets/SeattleHousing.csv') %>% 
  mutate(more_than2baths = bathrooms > 2)

z <- sum(seattle$more_than2baths)
N <- nrow(seattle)
```

#### a. (2 pts)
Justify your prior distribution.

*I have little knowledge of this parameter, and hence, will opt to use a uniform prior (Beta(1,1)). With a large data set, this will have minimal impact on the posterior distribution, relative to the data.*

#### b. (2 pts)
State the probability model you will using. You can, but don't need, to write out the full functional form of the probability mass/distribution function.

*We assume each house results in a Bernoulli trial with a probability term corresponding to having more than 2 bathrooms. Or alternatively, we could consider the collection of houses using a binomial distribution.*

#### c. (2 pts)
What is the form of your posterior distribution?

*Given this prior distribution and probability model for our data, the resultant posterior distribution is also a beta distribution, with parameters `r 1 + z` and `r 1 + N - z`*

#### d. (2 pts)
Plot your prior and posterior distributions on the same figure.

```{r}
num_x <- 100
x_seq <- seq(0,1, length.out = num_x)

tibble(dens = c(dbeta(x_seq, 1 + z, 1 + N - z),dbeta(x_seq, 1, 1)),
       x = c(x_seq, x_seq),
       type = c(rep('posterior', num_x), rep('prior', num_x))) %>%
  ggplot(aes(y=dens, x = x, color = type)) + 
  geom_line() + theme_bw() + theme(legend.position = 'bottom')

```


#### e. (2 pts)
Pretend your cousin has recently accepted a new job that requires relocating to Seattle. Summarize your findings (with regard to probability of finding a house with more than 2 bathrooms) in a non-technical manner avoiding statistical lingo.

*Hello Hans,*

*Congratulations on the new job - and the new baby girl. You are going to need extra bathrooms when those kids become teenagers. I did a quick analysis and found that you should be able to find a place with more than two bedrooms, roughly `r percent(qbeta(.025, 1 + z, 1 + N -z))` to `r percent(qbeta(.975, 1 + z, 1 + N -z))` of houses have more than 2 bathrooms. On the other hand, I hope the job pays well, because Seattle is very expensive.*